---
title: "Level 5 Bottle Data"
author: "Thomas Hutchinson"
date: "3/3/2021"
output: 
  pdf_document:
    latex_engine: xelatex
    dev: cairo_pdf
header-includes:
  - \usepackage{fontspec}
  - \setmainfont{IBM Plex Serif Light}
  - \setmonofont{IBM Plex Mono Light}
---
### Load packages and set environment

```{r message=FALSE}

library(tidyverse)
library(inspectdf)
library(tidymodels)
library(hrbrthemes)

```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Load Level 5 csv file

```{r message=FALSE}

data <- read_csv("level-5.csv")

```

Take out the metadata on the far right
```{r}

# Take out the metadata at the end of the table
data <- data %>% 
  select( -"Sample_Date",-"Substrate",-"Weeks",-"Sample_Code",
          -"Submerged",-"Floating",-"Water",-"Vegetation",
          -"Sediment",-"Cut",-"Shredded") %>% 
  pivot_longer(!index, names_to = "otu", values_to = "count")

# Take out the 0 value rows
data <- data %>% 
  filter(count>0)

```

Filter out chloropoasts
```{r}

data <- data %>% 
  filter(!grepl("Chloroplast", otu))

```



# Split into five columns
```{r}

t <-data

t <- t %>% separate(otu, c("d", "p", "c", "o", "f"), ";")

# t %>% select(p) %>% group_by(p) %>%  summarise(count=n())

```

# Take off prefixes

```{r}

t <- t %>% 
  select(index, d, p, count) %>% 
  filter(p != "__") %>% 
  filter(d != "__") %>% 
  drop_na() %>% 
  mutate(domain = str_sub(d,4),
         phylum = str_sub(p,4)) %>% 
  select(index, domain, phylum, count)

```

Sum by index and phylum.

```{r}

t <- t %>% 
  group_by(index, phylum) %>% 
  mutate(count = sum(count)) %>% 
  distinct()

```


```{r}

t <- t %>% 
  separate(index, c("sample_date", "sample_id"), "-")

t

```


filter out the four we want

```{r}

bottle <- t %>% 
  filter(sample_id %in% c("T06", "T07", "B16", "B17" )) 

```

Make a list of the top 10 across these four

```{r}

p_list <- t %>% 
  group_by(phylum) %>% 
  mutate(count = sum(count)) %>% 
  arrange(desc(count)) %>% 
  select(phylum) %>% 
  mutate(phylum = as.factor(phylum)) %>% 
  distinct() %>% 
  head(10)

p_list

```

Filter for just the ones on this list

```{r}

bottle <- p_list %>% 
  inner_join(y=t) %>% 
  select(sample_id, phylum, count) %>% 
  group_by(sample_id,phylum) %>% 
  mutate(count = sum(count)) %>% 
  distinct() %>% 
  arrange(sample_id) %>% 
  print(n=Inf)


```


Test to see it work with just one sample_id

```{r}

T06 <- bottle %>% 
  filter(sample_id == "T06") %>% 
  select(sample_id, phylum, count) %>% 
  group_by(phylum) %>% 
  mutate(count = sum(count)) %>% 
  distinct() %>% 
  arrange(desc(count)) %>% 
  head(10)

T07 <- bottle %>% 
  filter(sample_id == "T07") %>% 
  select(sample_id, phylum, count) %>% 
  group_by(phylum) %>% 
  mutate(count = sum(count)) %>% 
  distinct() %>% 
  arrange(desc(count)) %>% 
  head(10)

B16 <- bottle %>% 
  filter(sample_id == "B16") %>% 
  select(sample_id, phylum, count) %>% 
  group_by(phylum) %>% 
  mutate(count = sum(count)) %>% 
  distinct() %>% 
  arrange(desc(count)) %>% 
  head(10)

B17 <- bottle %>% 
  filter(sample_id == "B17") %>% 
  select(sample_id, phylum, count) %>% 
  group_by(phylum) %>% 
  mutate(count = sum(count)) %>% 
  distinct() %>% 
  arrange(desc(count)) %>% 
  head(10)

```


Combine into one tibble

```{r}

bottle <- bind_rows(T06, T07, B16, B17)

bottle

```


Make the horizontal plot

```{r}

bottle %>% 
  ggplot( aes(x=sample_id,
              y=count,
              fill = phylum) ) +
  geom_bar(position = "fill", 
           stat="identity") +
  theme_ipsum_ps() + 
  coord_flip()

```



Make the matrix

```{r}

cs_bottle <- as.matrix( bottle %>% 
  pivot_wider(names_from = sample_id, values_from = count) )

```


Chi-square

```{r}


chisq.test(cs_bottle)

```

